Bringing the Power of eBPF to Open vSwitch

### 2018-10-25 paper.tex
Edit ch 4.3.1

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn}.
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and provide a high speed channel from the XDP
directly to a userspace socket interface. In other word, AF\_XDP socket family
connects the XDP packet receiving/sending path to the userspace, by-passing the
rest of the Linux networking stacks.
An AF\_XDP socket, called XSK, is created by using the normal socket() system
call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet
buffer as parameter, XSK introduces two rings in userspace: the Rx ring and the
Tx ring. These two rings are per-XSK data structure that the userspace program
needs to properly configure and maintain in order to receive and send packets.
Moreover, to provide zerocopy support for AF\_XDP, all the packet data buffers
used in Tx/Rx rings are allocated from a specific memory region called \em{umem}
which consists of a number of fixed size memory chunks.
Two rings, the Fill ring and the Completion ring, are associated with a umem,
and a umem can be shared between multiple XSKs.
Each element in the Rx/Tx ring or in Fill/Completion ring is a descriptor that
contains an address that points to a chunk in umem. The address is not system's
virtual or physical address but simply an offset within the entire umem memory
region.

@@

\subsection{umem memory management}
In order to properly program Rx/Tx/Fill/Completion rings, we implement a
umem memory management layer, call umempool. Basically, it's a data structure
maintaining the unused/available elements in umem, and GET and PUT access APIs.
\subsubsection{Packet Reception}
\begin{figure}
{\scriptsize
\begin{verbatim}
umem: total 8 elements, each has 2K chunk size
                 +-------------------------------+
                 |   |   |   |   |   |   |   |   |
                 +-------------------------------+
            addr:  1   2   3   4   5   6   7  8 
(1) Initial Stat
           -------------------          umempool
Fill r   ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
           -------------------
           -------------------
Rx ring  ...|   |   |   |   | ...
           -------------------
(2) 4 Packets Arrive
           -------------------          umempool
Fill r   ...|   |   |   |   | ...       {5, 6, 7, 8}
          --------------------
           -------------------
Rx ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(3) Get 4 free umem elemet and refill
           -------------------          umempool
Fill r   ...| 5 | 6 | 7 | 8 | ...       {} 
           -------------------
           -------------------
Rx ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(4) Pass packets {1, 2, 3, 4} to dpif-netdev
    for parse, lookup, and actions.

@@

Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
the Rx ring for receving packets from XSK.  For simplicity to demonstrate
how AF\_XDP works, assume that there are only eight umem buffers, with each buffer's
size equals 2KB (?).  Initially at step 1, ovs-vswitchd fills four available umem elements into
the Fill ring and waits for incoming packets.  When there are incoming packets,
at step 2, the Fill ring's four buffer elems have been consumed and moved
to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
four available umem elements from the umempool, and fills into Fill ring (step 3).
Then ovs-vswitchd creates the metadata needed for the four packet
buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the umem buffer, a recycle
mechanism is triggered to place this buffer back to umempool (step 5).
Step 5 makes sure that there are always available elements in Fill ring, so
that the underlying XDP program in kernel can keep processing packets 
while the userspace ovs-vswitchd is processing the previous
received packets on Rx ring.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
new packets.

>>>>

AF\_XDP was proposed and upstreamed to Linux kernel since 4.18~\cite{af_xdp_lwn}.
The core idea behind the AF\_XDP is to leverage the XDP eBPF program's
early access to the raw packet, and provide a high speed channel from the XDP
directly to a userspace socket interface. In other word, AF\_XDP socket family
connects the XDP packet receiving/sending path to the userspace, by-passing the
rest of the Linux networking stacks.
An AF\_XDP socket, called XSK, is created by using the normal socket() system
call. Unlike AF\_PACKET which uses the send() and receive() syscalls with packet
buffer as parameter, XSK introduces two rings in userspace: the Rx ring and the
Tx ring. These two rings are per-XSK data structure that the userspace program
needs to properly configure and maintain in order to receive and send packets.
Moreover, to provide zerocopy support for AF\_XDP, all the packet data buffers
used in Tx/Rx rings are allocated from a specific memory region called {\em umem}
which consists of a number of fixed size memory chunks.
Two rings, the Fill ring and the Completion ring, are associated with a umem,
and a umem can be shared between multiple XSKs.
Each element in the Rx/Tx ring or in Fill/Completion ring is a descriptor that
contains an address that points to a chunk in umem. The address is not system's
virtual or physical address but simply an offset within the entire umem memory
region.

@@

\subsection{umem memory management}
In order to properly program Rx/Tx/Fill/Completion rings, we implement a
umem memory management layer, call umempool. Basically, it's a data structure
maintaining the unused/available elements in umem with GET and PUT access
APIs. We will demonstrate the use case of umempool with packet reception and
transmission in the following subsections.

\subsubsection{Packet Reception}
\begin{figure}
{\scriptsize
\begin{verbatim}
umem: total 8 elements, each has 2KB chunk size
                 +-------------------------------+
                 |   |   |   |   |   |   |   |   |
                 +-------------------------------+
            addr:  1   2   3   4   5   6   7  8 
(1) Initial Stat
           -------------------          umempool
Fill r   ...| 1 | 2 | 3 | 4 | ...       {5, 6, 7, 8}
           -------------------
           -------------------
Rx ring  ...|   |   |   |   | ...
           -------------------
(2) 4 Packets Arrive
           -------------------          umempool
Fill r   ...|   |   |   |   | ...       {5, 6, 7, 8}
          --------------------
           -------------------
Rx ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(3) Get 4 free umem elemet and refill
           -------------------          umempool
Fill r   ...| 5 | 6 | 7 | 8 | ...       {} 
           -------------------
           -------------------
Rx ring  ...| 1 | 2 | 3 | 4 | ...
           -------------------
(4) Pass packets {1, 2, 3, 4} to dpif-netdev
    for parse, lookup, and actions.

@@

Figure~\ref{afxdprx} shows how ovs-vswitchd sets up the Fill ring and
the Rx ring for receiving packets from XSK.  For simplicity, in this example,
we assume that there are only eight umem buffers, and each buffer's
size is 2KB.  Initially, at step 1, ovs-vswitchd pushes four available umem
elements into
the Fill ring and waits for incoming packets.  When there are incoming packets,
at step 2, the Fill ring's four buffer elements have been consumed and moved
to the Rx ring.  In order to keep receiving packets, ovs-vswitchd gets another
four available umem elements from the umempool, and fills into Fill ring (step 3).
Then, ovs-vswitchd creates the metadata needed for the four packet
buffer \{1, 2, 3, 4\}, i.e., struct dp\_packet and struct dp\_packet\_batch, and passes to 
the dpif-netdev layer for parse, lookup and action executions.
Finally, when ovs-vswitchd finishes processing the umem buffer, a recycle
mechanism is triggered to place this buffer back to umempool (step 5).
Step 5 makes sure that there are always available elements in Fill ring, so
that the underlying XDP program in kernel can keep processing packets 
while the userspace ovs-vswitchd is processing the previous
received packets on Rx ring.
When step 5 finishes, ovs-vswitchd goes back to step 2, waiting for
new packets.
